---
title: 'ANLY 510-90: Team GitHub Assignment'
author:
- Team SAS
- Sweta Yadav, Abhishek Chowdhary, Sathya Narayanan Subbaraj
- Harrisburg University
output:
  html_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80),tidy=TRUE, warning = FALSE)

```
<br>

## Data Cleanup
<br>
<br>

**Load Libraries **

```{r}

# Load required packages
library(stringr)
library(stringi)
library(zoo)

```
<br>
<br>


### Step-1)
<br>
**Coding done by: SWETA YADAV **
<br>
<br>

In step-1, the following actions were performed in R for data cleanup;
<br>
<br>
a) The dataset 'dirty_data.csv' was loaded into R.
<br>
b) The Area column with NA were populated by replacing with previous not NULL value.
<br>
c) The special characters were removed using 'gsub' function in R.
<br>
d) The variables 'Street' and 'Street.2' were trimmed for spaces at both ends and multiple spaces were replaced with single space.
<br>

```{r}

# Load dataset and set empty values to NA
dirty_data <- read.csv("dirty_data.csv", header = T, na.strings = c("","NA"))

# dirty_data column names
colnames(dirty_data)

# Populate the Area column by replacing NA with previous not NULL value
clean_data <- transform(dirty_data, Area = na.locf(Area))

# Remove special characters from column 'Street' and 'Street.2'
clean_data$Street <- gsub(pattern = "[^[:blank:][:alnum:]+.,&/\\-]", replacement = " ", clean_data$Street)
clean_data$Street.2 <- gsub(pattern = "[^[:blank:][:alnum:]+.,&/\\-]", replacement = " ", clean_data$Street.2)

# Trim leading and ending spaces, replace multiple spaces with single space
clean_data$Street <- gsub("\\s+", " ", str_trim(clean_data$Street))
clean_data$Street.2 <- gsub("\\s+", " ", str_trim(clean_data$Street.2))




```
<br>
<br>

