---
title: 'ANLY 510-90: Team GitHub Assignment'
author:
- Team SAS
- Sweta Yadav, Abhishek Chowdhary, Sathya Narayanan Subbaraj
- Harrisburg University
output:
  html_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80),tidy=TRUE, warning = FALSE)

```

## Data Cleanup

```{r}

# Load required packages
library(stringr)
library(stringi)
library(zoo)

```
### Coding done by: SWETA YADAV
#### Step-1) 
In step-1, the following actions were performed in R for data cleanup;


a) The dataset 'dirty_data.csv' was loaded into R.

b) The Area column with NA were populated by replacing with previous not NULL value.

c) The special characters were removed using 'gsub' function in R.

d) The variables 'Street' and 'Street.2' were trimmed for spaces at both ends and multiple spaces were replaced with single space.

```{r}

# Load dataset and set empty values to NA
dirty_data <- read.csv("dirty_data.csv", header = T, na.strings = c("","NA"))

# dirty_data column names
colnames(dirty_data)

# Populate the Area column by replacing NA with previous not NULL value
clean_data <- transform(dirty_data, Area = na.locf(Area))

# Remove special characters from column 'Street' and 'Street.2'
clean_data$Street <- gsub(pattern = "[^[:blank:][:alnum:]+,&/\\-]", replacement = " ", clean_data$Street)
clean_data$Street.2 <- gsub(pattern = "[^[:blank:][:alnum:]+,&/\\-]", replacement = " ", clean_data$Street.2)

# Trim leading and ending spaces, replace multiple spaces with single space
clean_data$Street <- gsub("\\s+", " ", str_trim(clean_data$Street))
clean_data$Street.2 <- gsub("\\s+", " ", str_trim(clean_data$Street.2))

```


### Coding done by: SATHYA NARAYANAN SUBBARAJ
#### Step-2)
In step-2, the following actions were performed in R for data cleanup;

a) The street text values in 'Street' and 'Street.2' were capitalized using Title case.

b) The street address denominations for Street, Avenue, Roads, Lane were standaridized using str_replace_all function.

c) Few other corrections to the 'Street' and 'Street.2' values were made.

```{r }

# Capitalization of the first letters of text values in 'Street' and 'Street.2'
clean_data$Street <- str_to_title(clean_data$Street)
clean_data$Street.2 <- str_to_title(clean_data$Street.2)

# Convert Street to St
clean_data$Street <- str_replace_all(clean_data$Street,"Street", "St")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Street", "St")

# Convert Road to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Road", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Road", "Rd")

# Convert Roads to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Roads", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Roads", "Rd")

# Convert Raod to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Raod", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Raod", "Rd")

# Convert Rds to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Rds", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Rds", "Rd")

# Convert Lane to Ln
clean_data$Street <- str_replace_all(clean_data$Street,"Lane", "Ln")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Lane", "Ln")

# Convert Avenue to Ave
clean_data$Street <- str_replace_all(clean_data$Street,"Avenue", "Ave")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Avenue", "Ave")


# Other Corrections
clean_data$Street <- gsub(pattern = " ,", replacement = ",", clean_data$Street)
clean_data$Street.2 <- gsub(pattern = " ,", replacement = ",", clean_data$Street.2)

clean_data$Street <- gsub(pattern = ",  ", replacement = ", ", clean_data$Street)
clean_data$Street.2 <- gsub(pattern = ",  ", replacement = ", ", clean_data$Street.2)

```

### Coding done by: ABHISHEK CHOWDHARY
#### Step-3) 
In step-3, the following actions were performed in R for data cleanup;

a) If the value in Street 2 duplicates the value in Street 1, the Street.2 value was removed using ifelse() function in R.

b) The Strange.HTML column was removed using subset function.

c) The final Cleaned up dataset was then exported.

```{r}

# If the value in Street 2 duplicates the value in Street 1, remove the value in Street 2
clean_data$Street.2 <- ifelse(clean_data$Street.2==clean_data$Street,"",clean_data$Street.2)

# Remove 'Strange.HTML' column by subsetting
final_data <- subset(clean_data, select = -Strange.HTML)

# Export the cleaned output
write.csv(final_data, file = "final_data.csv", row.names = FALSE, na = "")

```
