---
title: 'ANLY 510-90: Team GitHub Assignment'
author:
- Team SAS
- Sweta Yadav, Abhishek Chowdhary, Sathya Narayanan Subbaraj
- Harrisburg University
output:
  html_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80),tidy=TRUE, warning = FALSE)

```
<br>

## Data Cleanup
<br>
<br>

###Load Libraries

```{r}

# Load required packages
library(stringr)
library(stringi)
library(zoo)

```
<br>
<br>


### Step-1)
<br>
####Coding done by: SWETA YADAV
<br>
<br>

In step-1, the following actions were performed in R for data cleanup;
<br>
<br>
a) The dataset 'dirty_data.csv' was loaded into R.
<br>
b) The Area column with NA were populated by replacing with previous not NULL value.
<br>
c) The special characters were removed using 'gsub' function in R.
<br>
d) The variables 'Street' and 'Street.2' were trimmed for spaces at both ends and multiple spaces were replaced with single space.
<br>

```{r}

# Load dataset and set empty values to NA
dirty_data <- read.csv("dirty_data.csv", header = T, na.strings = c("","NA"))

# dirty_data column names
colnames(dirty_data)

# Populate the Area column by replacing NA with previous not NULL value
clean_data <- transform(dirty_data, Area = na.locf(Area))

# Remove special characters from column 'Street' and 'Street.2'
clean_data$Street <- gsub(pattern = "[^[:blank:][:alnum:]+,&/\\-]", replacement = " ", clean_data$Street)
clean_data$Street.2 <- gsub(pattern = "[^[:blank:][:alnum:]+,&/\\-]", replacement = " ", clean_data$Street.2)

# Trim leading and ending spaces, replace multiple spaces with single space
clean_data$Street <- gsub("\\s+", " ", str_trim(clean_data$Street))
clean_data$Street.2 <- gsub("\\s+", " ", str_trim(clean_data$Street.2))




```
<br>
<br>


### Step-2)
<br>
####Coding done by: SATHYA NARAYANAN SUBBARAJ
<br>
<br>

In step-2, the following actions were performed in R for data cleanup;
<br>
<br>
a) The street text values in 'Street' and 'Street.2' were capitalized using Title case.
<br>
b) The street address denominations for Street, Avenue, Roads, Lane were standaridized using str_replace_all function.
<br>
e) Few other corrections to the 'Street' and 'Street.2' values were made.

<br>
```{r}

# Capitalization of the first letters of text values in 'Street' and 'Street.2'
clean_data$Street <- str_to_title(clean_data$Street)
clean_data$Street.2 <- str_to_title(clean_data$Street.2)

# Convert Street to St
clean_data$Street <- str_replace_all(clean_data$Street,"Street", "St")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Street", "St")

# Convert Road to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Road", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Road", "Rd")

# Convert Roads to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Roads", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Roads", "Rd")

# Convert Raod to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Raod", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Raod", "Rd")

# Convert Rds to Rd
clean_data$Street <- str_replace_all(clean_data$Street,"Rds", "Rd")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Rds", "Rd")

# Convert Lane to Ln
clean_data$Street <- str_replace_all(clean_data$Street,"Lane", "Ln")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Lane", "Ln")

# Convert Avenue to Ave
clean_data$Street <- str_replace_all(clean_data$Street,"Avenue", "Ave")
clean_data$Street.2 <- str_replace_all(clean_data$Street.2,"Avenue", "Ave")


# Other Corrections
clean_data$Street <- gsub(pattern = " ,", replacement = ",", clean_data$Street)
clean_data$Street.2 <- gsub(pattern = " ,", replacement = ",", clean_data$Street.2)

clean_data$Street <- gsub(pattern = ",  ", replacement = ", ", clean_data$Street)
clean_data$Street.2 <- gsub(pattern = ",  ", replacement = ", ", clean_data$Street.2)

```
<br>
<br>

### Step-3)
<br>
####Coding done by: ABHISHEK CHOWDHARY
<br>
<br>

In step-3, the following actions were performed in R for data cleanup;
<br>
<br>
a) If the value in Street 2 duplicates the value in Street 1, the Street.2 value was removed using ifelse() function in R.
<br>
b) The Strange.HTML column was removed using subset function.
<br>
c) The final Cleaned up dataset was then exported.
<br>

```{r}

# If the value in Street 2 duplicates the value in Street 1, remove the value in Street 2
clean_data$Street.2 <- ifelse(clean_data$Street.2==clean_data$Street,"",clean_data$Street.2)

# Remove 'Strange.HTML' column by subsetting
final_data <- subset(clean_data, select = -Strange.HTML)

# Export the cleaned output
write.csv(final_data, file = "final_data.csv", row.names = FALSE, na = "")

```
<br>
<br>

